Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence

### 2019-04-24 eDriversMS-SI-v2.rmd
latest modifications including co-author comments and author contribution statement

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

@@

We used the annual Version 1 Nighttime VIIRS DNB composites between 2015 and
2016 [@eog2019] to characterize coastal development in coastal areas of the
St. Lawrence. As the effects of coastal development are likely acute in its
direct vicinity, we extracted average radiance values using a 2 km buffer
around grid cells within 2 km of the coast. We used a weighted area average to
extract the radiance values.

@@

As in @halpern2008 and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statscan2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008 and
@halpern2015b and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

@@

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008; @halpern2015] and available on the NCEAS
online data repository [@halpern2015b]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme [@jcomm2018]. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worlwide. Data used spanned 2003 to 2011.

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic [@tetreault2002]. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
was from November 2010 to December 2011.

@@

The data used to characterize invasive species risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008 and @halpern2015.

@@

The data used to characterize marine pollution risk comes from the global
cumulative impacts assessment on habitats [@halpern2008; @halpern2015] and
available on the NCEAS online data repository [@halpern2015b].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasives species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008 and @halpern2015.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015b] with our
1 $km^2$ grid cell using weighted area average.

>>>>

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

@@

We used the annual Version 1 Nighttime VIIRS DNB composites between 2015 and
2016 [@earthobservationgroup2019] to characterize coastal development in coastal areas of the
St. Lawrence. As the effects of coastal development are likely acute in its
direct vicinity, we extracted average radiance values using a 2 km buffer
around grid cells within 2 km of the coast. We used a weighted area average to
extract the radiance values.

@@

As in @halpern2008b and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statisticscanada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008b and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

@@

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008b; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worlwide. Data used spanned 2003 to 2011.

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
was from November 2010 to December 2011.

@@

The data used to characterize invasive species risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008b and @halpern2015.

@@

The data used to characterize marine pollution risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasives species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008b and @halpern2015.

For the St. Lawrence, we overlaid the raw data layers [@halpern2015] with our
1 $km^2$ grid cell using weighted area average.

### 2019-04-24 eDriversMS-v2.Rmd
latest modifications including co-author comments and author contribution statement

@@

The St. Lawrence System, formed by one of the largest estuaries in the world and
a vast interior sea, is a complex social-ecological system characterized by
highly variable environmental conditions and oceanographic processes, both in
space and time [@el-sabh1990; @white1997; @dufour2007]. It thus offers a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have historically
and contemporarily benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017 @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @goq2015; @rqm2018], an
intensification of the human footprint is expected in the St. Lawrence System.

As elsewhere in the world [see @halpern2015], this intensifying human footprint
will likely result in increasingly intricate environmental exposure regimes,
*i.e.* suites of overlapping and potentially interacting environmental drivers
threatening ecosystems, habitats or ecological communities. Drivers, often
referred to as stressors or pressures, are any externalities that affect
environmental processes and disturb natural systems. Drivers may originate from
natural or human-induced biophysical processes (*e.g.* sea surface temperature
anomalies and hypoxia) or directly from anthropogenic activities (*e.g.*
fisheries and marine pollution). The potential for complex interactions between
drivers is the largest uncertainty when studying or predicting environmental
change [@darling2008; @cote2016]. The effects of multiple drivers can combine
non-linearly and result in effects that are greater (synergistic effect) or
lower (antagonistic effect) than the sum of individual effects
[@crain2008; @darling2008; @cote2016].  <!-- For example...??? -->

@@

The current digital infrastructure is highly decentralized and the data
management and sharing practices are highly heterogeneous, preventing us from
maximizing benefits from research investments [@wilkinson2016].
Yet there now exists multiple initiatives that address this issue by assembling,
organizing and sharing vast arrays of environmental data.
Biotic data can be accessed through web portals such as the Ocean Biogeographic
Information System [OBIS; @obis2018], the Global Biodiversity Information Facility
[GBIF; @gbif2018], the Global Biotic Interactions platform [GloBI; @poelen2014],
and the World Register of Marine Species [WoRMS; @worms2017]. Abiotic data can
also be accessed through WorldClim [@hijmans2005], Bio-ORACLE
[@tyberghein2012], and MARSPEC [@sbrocco2013]. Initiatives focused on sharing
environmental data for specific areas also exist, such as the St. Lawrence
Global Observatory (SLGO; https://ogsl.ca/en), the European Marine Observation
and Data Network (EMODnet; http://www.emodnet.eu/) and the U.S. Integrated Ocean
Observing System (IOOS; https://ioos.noaa.gov/). Essential environmental
parameters are also organized, coordinated and acquired through global
initiatives like the Group on Earth Observations Biodiversity Observation
Network [GEO BON; @scholes2012], the Census of Marine Life [CoML; @coml2010]
and the Global Ocean Observing System (GOOS; http://www.goosocean.org).

@@

The list of drivers for which we sought data was informed by a global cumulative
impact assessment initiative [@halpern2008; @halpern2015], regional holistic
evaluations of the state of the St. Lawrence [@dufour2007; @benoit2012], and
communications with regional experts. Through the data gathering process, we
developed and continue to develop collaborations with regional experts and data
holders. We also use global data from the global cumulative impact assessment
initiative [@halpern2008; @halpern2015] available from the National Center
for Ecological Analysis and Synthesis (NCEAS) online data repository
[Table 1; @halpern2015b]. We selected global data that were unavailable at the
regional scale and that were available at a resolution adequate for use at the
scale of the St. Lawrence (*e.g.* marine pollution).

@@

These results are consistent with observations elsewhere in the world, where
cumulative driver exposure conspicuously arises from and markedly intensifies
close to coastal cities and at the mouth of rivers draining highly populated
areas [e.g. @halpern2015; @feist2016; @mach2017; @stock2018]. These are areas
where human activities (*e.g.* coastal development and shipping) and footprint
(*e.g.* pollution runoff) are the most intense [@feist2016], and on which is
overlaid a background of natural disturbances [@micheli2016]. They are also the
areas in which the most dramatic increases in exposure are expected, with
populations increasing more rapidly along the coast than inland [@feist2016].
In the St. Lawrence, large coastal cities are mostly located along the Estuary
and the southwestern Gulf, while the northeastern Gulf is largely uninhabited
or home to small coastal communities.

As for offshore exposure, the Estuary, along with the St. Lawrence River,
provide access to and serve as the primary drainage outflow of the Great Lakes
Basin, the most densely populated region in Canada [@statcan2017]. Most marine
traffic thus converges to the Estuary.

@@

Fisheries in the St. Lawrence have historically affected
biodiversity distribution and habitat quality [@moritz2015]. Concurrently,
hypoxia decreases overall habitat quality, but triggers species-dependent
responses ranging from adaptation [*e.g.* northern shrimp *Pandalus borealis*
and Greenland halibut *Reinhardtius hippoglossoides*; @pillet2016] to reduced
growth rates [@dupont2013] and avoidance of oxygen-depleted habitats
[*e.g.* Atlantic cod *Gadus morhua*; @chabot2008] to increased mortality
[*e.g.* sessile benthic invertebrates; @eby2005; @gilbert2007; @belley2010].
Certain species may thus be adversely affected by fisheries and withstand
hypoxia but still experience a decrease in prey availability, while others may
be deleteriously affected by the compounded effect of both drivers [@deleo2017].

@@

***How***: By promoting, consolidating and working with experts involved in
existing and highly valuable environmental initiatives already in place in the
St. Lawrence. Notable examples of environmental initiatives are the annual
review of physical [@galbraith2018], chemical, and biological [@blais2018]
oceanographic conditions in the St. Lawrence, the fisheries monitoring program
[@dfo2016], the annual groundfish and shrimp multidisciplinary survey
[@bourdages2018], the characterisation of benthic [@dutil2011], epipelagic and
coastal [@dutil2012] habitats of the St. Lawrence, and Canada's shoreline
classification [@eccc2018]. There are also nascent efforts to share
information on several human activities in the St. Lawrence such as the Marine
Spatial Data Infrastructure portal, which provides data on zoning, shipping,
port activities, and other human activities in Canadian waters, including the
St. Lawrence system [@goc2016].

By working with existing data portals whose objective is to share
environmental data. We are thus collaborating actively with the St. Lawrence
Global Observatory (SLGO) to develop the initiative and to host the
platform on their web portal (https://ogsl.ca/en). The mission of SLGO is to
promote and facilitate the accessibility, dissemination and exchange of official
and quality data and information on the St. Lawrence ecosystem through the
networking of organisations and data holders to meet their needs and those of
users, to improve knowledge and to assist decision-making in areas such as
public safety, climate change, transportation, resources and biodiversity
conservation. The SLGO is also one of three regional association spearheading
the Canadian Integrated Ocean Observing System
(CIOOS; http://meopar.ca/research/cioos-call-for-proposals/),
which will focus on integrating oceanographic data from multiple sources to
make them accessible to end-users and to enable the national coordination of
ocean observing efforts by integrating isolated or inaccessible data, and by
identifying gaps or duplications in observations and research efforts.
We are also developing collaborations with the Portal on water knowledge
(http://www.environnement.gouv.qc.ca/eau/portail/), an initiative from the
Québec provincial government. This portal aims at collecting and sharing
accurate, complete and updated ressources on water and aquatic ecosystems to
support the mandate of relevant actors and stakeholders working in
water and aquatic ecosystems management.

@@

By inviting external community contributions (Figure \ref{diag}).
External researchers or entities wishing to submit marine data will be able to
do so through the SLGO web portal (https://ogsl.ca/en). Submissions through
other data portals will also be accepted either through the development of data
sharing agreements or with the caveat that shared data are under an open source
license and that they adhere to the platform data standards.

@@

***Why***: Open data has been propelled to the forefront of scientific research
in an era of open, collaborative and reproducible science. By moving towards
large scale, cross-disciplinary research and management projects, there is a
growing need to increase the efficiency of data discovery, access,
interoperability and analysis [@reichman2011; @wilkinson2016]. Our goal is to
foster efficient and functional open science by creating a fully open,
transparent and replicable open knowledge platform.

@@

By defining clear data and metadata standards and specifications to support
the regional standardization of current and future protocols and practices and
to favour interoperability with national and international initiatives like the
Essential Ocean Variables (EOV) identified by the Global Ocean Observing System
[GOOS; http://www.goosocean.org]. As such, we will adopt the metadata standard
currently targetted for the upcoming CIOOS, *i.e.* the North American Profile
of ISO 19115:2014 - Geographic information - Metadata, a schema favoured
for geospatial data in Canada and the United-States.

@@

***Why***: In the face of uncertainty and in an effort to address impending
environmental changes, adaptive management has been identified as the chief
strategy to guide efficient decision-making
[*e.g.* @costanza1998; @margules2000; @keith2011; @jones2016] and has already
been discussed in the context of multi-drivers and cumulative impact assessments
[@halpern2015; @beauchesne2016; @cote2016; @schloss2017]. Adaptive management can only be
truly achieved through a commitment to adaptive monitoring and data reporting
[@margules2000; @halpern2012; @lubchenco2015]. We further contend that adaptive
management requires the development of adaptive monitoring tools and
infrastructures, which we seek to address through a continuously-evolving platform.

@@

Critically - and we emphasize this point - *eDrivers* will
allow the scientific and governmental communities to identify key knowledge
gaps that will assist in prioritizing and optimizing research efforts.
Ultimately, we believe that *eDrivers* will operationalize evidence-based
decision-making by streamlining data management and research, allowing science
output to be available and interpretable on a time scale relevant to management
[see @reichman2011]. The platform will thus greatly facilitate the application
of broad scale, holistic research and management approaches such as
ecosystem-based management, strategic environmental assessments and
social-ecological metanetworks [*e.g.* @halpern2015; @jones2016; @dee2017].

@@

TO WRITE

>>>>

**Keywords**: ocean observing systems, St. Lawrence, environmental drivers,
cumulative exposure, threat complex, multiple stressors, global change

@@

The St. Lawrence System, formed by one of the largest estuaries in the world and
a vast interior sea, is a complex social-ecological system characterized by
highly variable environmental conditions and oceanographic processes, both in
space and time [@el-sabh1990a; @white1997; @dufour2007]. It thus offers a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have historically
and contemporarily benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017 @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @governmentofquebec2015; @rqm2018], an
intensification of the human footprint is expected in the St. Lawrence System.

As elsewhere in the world [see @halpern2015a], this intensifying human footprint
will likely result in increasingly intricate environmental exposure regimes,
*i.e.* suites of overlapping and potentially interacting environmental drivers
threatening ecosystems, habitats or ecological communities. Drivers, often
referred to as stressors or pressures, are any externalities that affect
environmental processes and disturb natural systems. Drivers may originate from
natural or human-induced biophysical processes (*e.g.* sea surface temperature
anomalies and hypoxia) or directly from anthropogenic activities (*e.g.*
fisheries and marine pollution). The potential for complex interactions between
drivers is the largest uncertainty when studying or predicting environmental
change [@darling2008; @cote2016]. The effects of multiple drivers can combine
non-linearly and result in effects that are greater (synergistic effect) or
lower (antagonistic effect) than the sum of individual effects
[@crain2008; @darling2008; @cote2016].  <!-- For example...??? -->

@@

The current digital infrastructure is highly decentralized and the data
management and sharing practices are highly heterogeneous, preventing us from
maximizing benefits from research investments [@wilkinson2016].
Yet there now exists multiple initiatives that address this issue by assembling,
organizing and sharing vast arrays of environmental data.
Biotic data can be accessed through web portals such as the Ocean Biogeographic
Information System [OBIS; @obis2018], the Global Biodiversity Information Facility
[GBIF; @gbif2018], the Global Biotic Interactions platform [GloBI; @poelen2014],
and the World Register of Marine Species [WoRMS; @wormseditorialboard2017]. Abiotic data can
also be accessed through WorldClim [@hijmans2005], Bio-ORACLE
[@tyberghein2012], and MARSPEC [@sbrocco2013]. Initiatives focused on sharing
environmental data for specific areas also exist, such as the St. Lawrence
Global Observatory (SLGO; https://ogsl.ca/en), the European Marine Observation
and Data Network (EMODnet; http://www.emodnet.eu/) and the U.S. Integrated Ocean
Observing System (IOOS; https://ioos.noaa.gov/). Essential environmental
parameters are also organized, coordinated and acquired through global
initiatives like the Group on Earth Observations Biodiversity Observation
Network [GEO BON; @scholes2012], the Census of Marine Life [CoML; @coml2010]
and the Global Ocean Observing System (GOOS; http://www.goosocean.org).

@@

The list of drivers for which we sought data was informed by a global cumulative
impact assessment initiative [@halpern2008b; @halpern2015a], regional holistic
evaluations of the state of the St. Lawrence [@dufour2007; @benoit2012], and
communications with regional experts. Through the data gathering process, we
developed and continue to develop collaborations with regional experts and data
holders. We also use global data from the global cumulative impact assessment
initiative [@halpern2008b; @halpern2015a] available from the National Center
for Ecological Analysis and Synthesis (NCEAS) online data repository
[Table 1; @halpern2015]. We selected global data that were unavailable at the
regional scale and that were available at a resolution adequate for use at the
scale of the St. Lawrence (*e.g.* marine pollution).

@@

These results are consistent with observations elsewhere in the world, where
cumulative driver exposure conspicuously arises from and markedly intensifies
close to coastal cities and at the mouth of rivers draining highly populated
areas [e.g. @halpern2015a; @feist2016; @mach2017; @stock2018]. These are areas
where human activities (*e.g.* coastal development and shipping) and footprint
(*e.g.* pollution runoff) are the most intense [@feist2016], and on which is
overlaid a background of natural disturbances [@micheli2016]. They are also the
areas in which the most dramatic increases in exposure are expected, with
populations increasing more rapidly along the coast than inland [@feist2016].
In the St. Lawrence, large coastal cities are mostly located along the Estuary
and the southwestern Gulf, while the northeastern Gulf is largely uninhabited
or home to small coastal communities.

As for offshore exposure, the Estuary, along with the St. Lawrence River,
provide access to and serve as the primary drainage outflow of the Great Lakes
Basin, the most densely populated region in Canada [@statisticscanada2017]. Most marine
traffic thus converges to the Estuary.

@@

Fisheries in the St. Lawrence have historically affected
biodiversity distribution and habitat quality [@moritz2015]. Concurrently,
hypoxia decreases overall habitat quality, but triggers species-dependent
responses ranging from adaptation [*e.g.* northern shrimp *Pandalus borealis*
and Greenland halibut *Reinhardtius hippoglossoides*; @pillet2016] to reduced
growth rates [@dupont-prinet2013] and avoidance of oxygen-depleted habitats
[*e.g.* Atlantic cod *Gadus morhua*; @chabot2008] to increased mortality
[*e.g.* sessile benthic invertebrates; @eby2005; @gilbert2007; @belley2010].
Certain species may thus be adversely affected by fisheries and withstand
hypoxia but still experience a decrease in prey availability, while others may
be deleteriously affected by the compounded effect of both drivers [@deleo2017].

@@

***How***: By promoting, consolidating and working with experts involved in
existing and highly valuable environmental initiatives already in place in the
St. Lawrence. Notable examples of environmental initiatives are the annual
review of physical [@galbraith2018], chemical, and biological [@blais2018]
oceanographic conditions in the St. Lawrence, the fisheries monitoring program
[@dfo2016], the annual groundfish and shrimp multidisciplinary survey
[@bourdages2018], the characterisation of benthic [@dutil2011], epipelagic and
coastal [@dutil2012] habitats of the St. Lawrence, and Canada's shoreline
classification [@eccc2018]. There are also nascent efforts to share
information on several human activities in the St. Lawrence such as the Marine
Spatial Data Infrastructure portal, which provides data on zoning, shipping,
port activities, and other human activities in Canadian waters, including the
St. Lawrence system [@governmentofcanada2018].

By working with existing data portals whose objective is to share
environmental data. We are thus collaborating actively with the St. Lawrence
Global Observatory (SLGO) to develop the initiative and to host the
platform on their web portal (https://ogsl.ca/en). The mission of SLGO is to
promote and facilitate the accessibility, dissemination and exchange of official
and quality data and information on the St. Lawrence ecosystem through the
networking of organisations and data holders to meet their needs and those of
users, to improve knowledge and to assist decision-making in areas such as
public safety, climate change, transportation, resources and biodiversity
conservation. SLGO is also one of three regional associations spearheading
the Canadian Integrated Ocean Observing System (CIOOS; https://cioos.ca),
which will focus on integrating oceanographic data from multiple sources to
make them accessible to end-users and to enable the national coordination of
ocean observing efforts by integrating isolated or inaccessible data, and by
identifying gaps or duplications in observations and research efforts.
We are also developing collaborations with the Portal on water knowledge
(http://www.environnement.gouv.qc.ca/eau/portail/), an initiative from the
Québec provincial government. This portal aims at collecting and sharing
accurate, complete and updated ressources on water and aquatic ecosystems to
support the mandate of relevant actors and stakeholders working in
water and aquatic ecosystems management.

@@

By inviting external community contributions (Figure \ref{diag}).
External researchers or entities wishing to submit marine data will be able to
do so through SLGO web portal (https://ogsl.ca/en). Submissions through
other data portals will also be accepted either through the development of data
sharing agreements or with the caveat that shared data are under an open source
license and that they adhere to the platform data standards.

@@

***Why***: Open knowledge has been propelled to the forefront of scientific research
in an era of open, collaborative and reproducible science. By moving towards
large scale, cross-disciplinary research and management projects, there is a
growing need to increase the efficiency of data discovery, access,
interoperability and analysis [@reichman2011; @wilkinson2016]. Our goal is to
foster efficient and functional open science by creating a fully open,
transparent and replicable open knowledge platform.

@@

By defining clear data and metadata standards and specifications to support
the regional standardization of current and future protocols and practices and
to favour interoperability with national and international initiatives like the
Essential Ocean Variables (EOV) identified by the Global Ocean Observing System
[GOOS; http://www.goosocean.org]. As such, we will adopt the metadata standard
currently targetted for CIOOS, *i.e.* the North American Profile
of ISO 19115:2014 - Geographic information - Metadata, a schema favoured
for geospatial data in Canada and the United States.

@@

***Why***: In the face of uncertainty and in an effort to address impending
environmental changes, adaptive management has been identified as the chief
strategy to guide efficient decision-making
[*e.g.* @margules2000; @keith2011; @jones2016; @chion2018] and has already
been discussed in the context of multi-drivers and cumulative impact assessments
[@halpern2015a; @beauchesne2016; @cote2016; @schloss2017]. Adaptive management can only be
truly achieved through a commitment to adaptive monitoring and data reporting
[@margules2000; @halpern2012; @lubchenco2015]. We further contend that adaptive
management requires the development of adaptive monitoring tools and
infrastructures, which we seek to address through a continuously-evolving platform.

@@

Critically - and we emphasize this point - *eDrivers* will
allow the scientific and governmental communities to identify key knowledge
gaps that will assist in prioritizing and optimizing research efforts.
Ultimately, we believe that *eDrivers* will operationalize evidence-based
decision-making by streamlining data management and research, allowing science
output to be available and interpretable on a time scale relevant to management
[see @sutherland2004; @reichman2011]. The platform will thus greatly facilitate the application
of broad scale, holistic research and management approaches such as
ecosystem-based management, strategic environmental assessments and
social-ecological metanetworks [*e.g.* @halpern2015a; @jones2016; @dee2017].

@@

DB, RD, DG and PA conceived the manuscript and the underlying objective. DB
prepared/formatted the data, performed the analyses, was in charge of technical
developments and lead the drafting of the manuscript. All co-authors contributed
to data, analases and writing based on their respective expertise and
contributed to the revision of the manuscript.
